---
title: "Capstone Sofware Development in R Vignette"
author: "Dong"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{EQVT Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This is an R package to analyse the data  ["NOAA Significant Earthquake Database"](https://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=1&d=1) in Capstone Sofware Development in R.


## Capstone's functions:
  
- `eq_clean_data` and `eq_location_clean` functions clean the NOAA data frame: 
- `eq_map` function maps the earthquakes epicenters (LATITUDE/LONGITUDE) with the desired annotation
- `eq_create_label` function that creates an HTML label for Location, Total deaths, and Magnitude in the leaflet map.
- Two geoms that visualize the NOAA earthquakes dataset. The `geom_timeline` function visualizes the times, the magnitudes and the number of deaths corresponding to earthquakes within a given countries respectively, while the `geom_timelinelabel` adds annotations to the earthquake .
     
## Examples

Download the data signif.txt from the NOAA website, put it to your working directory and transform it to a data frame using the `read_delim` function:
  
```{r eval = FALSE}
library(readr)
data <- readr::read_delim("signif.txt", delim = "\t")
```


Then clean the data with the functions `eq_clean_data` and `eq_location_clean`.

```{r eval = FALSE}
clean_data <- eq_clean_data(data)
clean_data <- eq_location_clean(clean_data)
```

To map the earthquakes epicenters in Canada since 2000 and get their dates in annotation use the `eq_map` function.

```{r eval = FALSE}
clean_data %>%
  dplyr::filter(COUNTRY == "CANADA" & lubridate::year(DATE) >= 2000) %>%
  eq_map(annot_col = "DATE")
```

To get annotation the Location, Total deaths, and Magnitude of the earthquakes, use the `eq_create_label` function before the `eq_map` function.
```{r eval = FALSE}
clean_data %>%
  dplyr::filter(COUNTRY == "CANADA" & lubridate::year(DATE) >= 2000) %>%
  dplyr::mutate(popup_text = eq_create_label(.)) %>%
  eq_map(annot_col = "popup_text")
```

To visualize times, magnitudes of earthquakes in Canada and USA, we use the `geom_timeline` geom.

```{r eval = FALSE}
data %>%
  dplyr::filter(COUNTRY == c("CANADA","USA") & lubridate::year(DATE) >= 2010) %>%
  ggplot(aes(x=DATE,y=COUNTRY,color=TOTAL_DEATHS,size=EQ_PRIMARY)) +
  geom_timeline(alpha=.5) +
  theme(legend.position="bottom", legend.box="horizontal", plot.title=element_text(hjust=0.5)) +
  ggtitle("Earthquakes") 
```

To add annotations to the earthquake data, we use `geom_timelinelabel` geom 

```{r eval = FALSE}
data %>%
  dplyr::filter(COUNTRY == c("CANADA","USA") & lubridate::year(DATE) >= 2010) %>%
  ggplot(aes(x=DATE,y=COUNTRY,color=TOTAL_DEATHS,size=EQ_PRIMARY)) +
  geom_timeline(alpha=.5) +
  geom_timelinelabel(aes(label=LOCATION_NAME),n_max=3) +
  theme(legend.position="bottom", legend.box="horizontal", plot.title=element_text(hjust=0.5)) +
  ggtitle("Earthquakes") 
```
